---
title: "It's About Time: Using Survival Analysis to Gain Time-Based People Insights"
author: "Behailu Bekera"
date: "April 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: House Keeping


```{r }
#Clear your workspace
#rm(list = ls()) 

#set working directory to where the data resides
setwd("C:/Users/bbekera/Desktop/SIOP Master Tutorial")
load(file = "Data/Survival.RData")
```

```{r message=FALSE}
#Load required packages
library(survival)
library(dplyr)
library(ranger)
library(ggplot2)
library(survminer)
library(pec)
```

```{r message=FALSE}
#Load data
data <- merge_data
```

Columns:

applicant_id  
term_reason_text  
application_date  
hire_date  
term_date  
emp1_start_date  
emp1_end_date  
new_job_title_hire  
times_available  
times_not_available  
highest_education  
overall_percentile  
storelocation  
home_state_province  
home_city  


# Part 1: Data preparation

```{r message=FALSE}
#Remove temporary positions and reduction in force
data <- subset(data, term_reason_text != "Invol-Elimination of Position")
data <- subset(data, term_reason_text != "Invol-End of Temporary Pos")
data <- subset(data, term_reason_text != "Invol - Location Closure")
```

```{r message=FALSE}

#Date variables to format into "%m/%d/%Y" format
datvar<-c("application_date","hire_date","term_date","emp1_start_date","emp1_end_date","emp2_start_date","emp2_end_date","emp3_start_date","emp3_end_date","emp4_start_date","emp4_end_date","edu1_start_date", "edu2_start_date","edu3_start_date","edu4_start_date")

#Format date variables
data[,c(which(colnames(data) %in% datvar))] <- lapply(data[c(which(colnames(data) %in% datvar))],function(x) as.Date(x,format="%m/%d/%Y"))
```

```{r message=FALSE}
#Calculate best guess of data extract date as the maximum termination date reported
#this is the right censor date for the survival analysis
extract_date <- max(data$term_date, na.rm=T)

#Calculate raw tenure variables
data$tenure_raw <- as.numeric(data$term_date - data$hire_date)

#counts how many cases have a hire_date after the term_date
sum(data$hire_date>data$term_date, na.rm=T) 

#Flag for excluding negative values for tenure
data$exclude<-ifelse(data$hire_date>data$term_date,1,0)
```
```{r message=FALSE}
#Used to substitute tenure value when term_date is NA
data$since_hire <- extract_date - data$hire_date

#Create clean continuous tenure variable
#This is the time to event 
data$tenure_cont <- NA

for(i in 1:nrow(data)){
  #minimum of raw tenure or days since hired; only fills in days since hire if the raw is NA
  data$tenure_cont[i] <- min(c(data$tenure_raw[i],data$since_hire[i]), na.rm=T) 
} 

#Overwrite NA for nonsense values
data$tenure_cont[data$exclude == 1] <- NA  

#Tenure summary stats
summary(data$tenure_cont)
```

```{r}
hist(data$tenure_cont, main = "Distribution of Time to Event", xlab = "Tenure_cont")
```




```{r message=FALSE}
#create attrition variable to indicate whether the event happended during the study period
#sum(is.na(data$term_date))
data$attrited <- ifelse(is.na(data$term_date) == TRUE,0,1)

#Number of attrites and those stayed
table(data$attrited)

#Tenure distribution among stayers 
summary(data$tenure_cont[data$attrited == 0])

#Tenure distribution among stayers
summary(data$tenure_cont[data$attrited == 1])
```





# Part 2: Survival Curve Analysis

```{r message=FALSE}
#Retain data for those currently with a Sales Associate job title
sa_data <- subset(data, new_job_title_hire == "Sales Associate")
```

```{r message=FALSE}
#Create a survival object that is used as a response variable in a model formula. 
sa_surv <- Surv(sa_data$tenure_cont, sa_data$attrited)

#Right-censored times to event (where attrition wasn't observed yet) have a "+" sign
#Printing out the first 30
head(sa_surv,30)

#Printing out tenure values for the first 30 
sa_data$tenure_cont[0:30]
```



The formula sa_surv ~ 1 instructs the survfit() function to fit a model with intercept only,   
and produces the Kaplan-Meier survival curve estimate.

```{r message=FALSE}
sa_km_fit <- survfit(sa_surv ~ 1)
```

```{r}
#Option 1 ploting KM curve
plot(sa_km_fit,
     main = 'Survival Rate for Sales Associates',ylab = "Survival Rate", xlab = "Days")
legend(1000, 0.99, legend = c('K-M survival estimate'), lty = 1:2)
```


```{r, echo=FALSE}
#Option 2 plotting using surviminer package
ggsurv <- ggsurvplot(
  sa_km_fit,                  # survfit object with calculated statistics.
  data = sa_data,             # data used to fit survival curves.
  palette = c("#32CD32"),
  xlim = c(0,700),         # present narrower X axis, but not affects.
  xlab = "Time in days",   # customize X axis label.
  ylab="Survival Rate",
  break.time.by = 100,     # break X axis in time intervals by 500.
  ggtheme = theme_light(), # customize plot and risk table with a theme.
  surv.median.line = "hv"  # add the median survival pointer.
  ,title='Survival Rate for Sales Associates'
  )
ggsurv
```

To get the number of days after which survival Rate reaches a certain percentage (e.g. 75%, 50%, 25%, 62%, 43%, 26%)

```{r}
#head(summary(sa_km_fit))

Time_and_prob=cbind.data.frame(sa_km_fit$time,sa_km_fit$surv)
colnames(Time_and_prob)<-c("time", "surv")

Time_and_prob$surv= round(Time_and_prob$surv, 2)

Probs= c(0.75, 0.5, 0.25, 0.62, 0.43, 0.26)

for (i in 1: length(Probs)){
  print (paste("Survival Rate reaches ", Probs[i], "%", " on day ", max(Time_and_prob$time[Time_and_prob$surv==Probs[i]]),sep = ""))
}
```

## Part 3: Feature Engineering

Creating features associated with availability:

```{r}
#See some of the raw times available values
head(data$times_available,1)

#Initialize new times available features
data$time_slots_available <- 0
data$weekends_only <- NA
data$weekdays_only <- NA
data$both <- NA

#Populate these times available features with values
for (i in 1:nrow(data)){
  temp <- as.character(data$times_available[i])
  
  #Number of time slots available
  data$time_slots_available[i] <- length(unlist(strsplit(temp,";")))
  
  #1 if available weekends ONLY, 0 otherwise even if no time slots mentioned at all
  data$weekends_only[i] <- ifelse(grepl("(Sat|Sun)",temp) == TRUE & grepl("(Mon|Tues|Wed|Thurs|Fri)",temp) == FALSE,1,0)
  
  #1 if available weekdays ONLY, 0 otherwise even if no time slots mentioned at all
  data$weekdays_only[i] <- ifelse(grepl("(Sat|Sun)",temp) == FALSE & grepl("(Mon|Tues|Wed|Thurs|Fri)",temp) == TRUE,1,0)
  
  #1 if available both weekdays and weekends, 0 otherwise even if no time slots mentioned at all
  data$both[i] <- ifelse(grepl("(Sat|Sun)",temp) == TRUE & grepl("(Mon|Tues|Wed|Thurs|Fri)",temp) == TRUE,1,0)
}

#Fill in NA if no time slots selected, that
data$weekends_only[data$time_slots_available == 0] <- NA # Now, data$weekends_only==1 => weekends_only and 0 means weekdays, NA mean no time slots given
data$weekdays_only[data$time_slots_available == 0] <- NA #Now, data$weekdays_only==1 => weekdays_only and 0 means weekends, NA mean no time slots given
data$both[data$time_slots_available == 0] <- NA # Now, data$both==1 => mentioned both weekends and weekdays, NA means neither(no time slots given)


#table(data$time_slots_available) #Number of time slots applicant is available for
#table(data$weekends_only)
#table(data$weekdays_only)
#table(data$both)
```


Creating Education Level features:

```{r}
#Highest education values
table(data$highest_education)

#Renaming and standardizing highest edudaction data
data$highest_education[data$highest_education == "Associates&#39;s Degree"] <- "Associates's Degree"
data$highest_education[data$highest_education == "Associatess Degree"] <- "Associates's Degree"
data$highest_education[data$highest_education == "Bachelor&#39;s Degree"] <- "Bachelor's Degree"
data$highest_education[data$highest_education == "Bachelors Degree"] <- "Bachelor's Degree"
data$highest_education[data$highest_education == "Master&#39;s Degree"] <- "Master's Degree"
data$highest_education[data$highest_education == "Masters Degree"] <- "Master's Degree"
data$highest_education <- factor(data$highest_education)
```

```{r}
#Creating binary versions of highest education features
data$high_school <- ifelse(data$highest_education == "High School or Equivalency",1,0)
data$associates <- ifelse(data$highest_education == "Associates's Degree",1,0)
data$bachelor <- ifelse(data$highest_education == "Bachelor's Degree",1,0)
data$masters_plus <- ifelse(data$highest_education == "Master's Degree" | data$highest_education == "Doctoral Degree",1,0)

data$college <- ifelse(data$highest_education == "Associates's Degree"|
                         data$highest_education == "Bachelor's Degree"|
                         data$highest_education == "College Diploma"|
                         data$highest_education == "Doctoral Degree"|
                         data$highest_education == "Master's Degree",1,0)
```

```{r}
#Number of applicants vs non
#table(data$high_school)
#table(data$associates)
#table(data$bachelor)
#table(data$masters_plus)
#table(data$college)
```



Creating rehire related feature

```{r, echo=FALSE}
data$rehire <- 0
data$rehire[grepl("Pier (One|1)", data$emp1_name, ignore.case = TRUE) == TRUE] <- 1
data$rehire[grepl("Pier (One|1)", data$emp2_name, ignore.case = TRUE) == TRUE] <- 1
data$rehire[grepl("Pier (One|1)", data$emp3_name, ignore.case = TRUE) == TRUE] <- 1

```

## Part 4: Feature-specific survival curve analysis

```{r}
#Subset sales associate data beause that is the role this model is going to be built for.
sa_data <- subset(data, new_job_title_hire == "Sales Associate")
sl_data <- subset(data, new_job_title_hire == "Sales Leader")
```


```{r}
#Overlay KM Plots based on job applied for

#Applied to Sales Associate role
sa_data_sa_app <- subset(sa_data, new_job_group_application == "Retail Salespersons")

#Applied to Sales Leader role
sa_data_sl_app <- subset(sa_data, new_job_group_application == "First-Line Supervisors of Retail Sales Workers")

#Create survival object for both types of applicants
sa_surv_sa_app <- Surv(sa_data_sa_app$tenure_cont, sa_data_sa_app$attrited)
sa_sa_app_km_fit <- survfit(sa_surv_sa_app ~ 1)

sa_surv_sl_app <- Surv(sa_data_sl_app$tenure_cont, sa_data_sl_app$attrited)
sa_sl_app_km_fit <- survfit(sa_surv_sl_app ~ 1)

plot(sa_sa_app_km_fit$surv,
     main = 'Position Applied Fore', ylab = "Survival Rate", xlab = "Days",lty = 2, pch = 20)
points(sa_sl_app_km_fit$surv, col = "red",lty = 2, pch = 20)
legend(400, 0.99, legend = c('Associate','Leader'),  
                        col = c("black","red"), pch = c(20,20),
                        lwd = c(1,1))

```

Rehire or not:

```{r}
wdata <- subset(sa_data, rehire == 1)
sa_surv_rehire_yes <- Surv(wdata$tenure_cont, wdata$attrited)
sa_rehire_yes_km_fit <- survfit(sa_surv_rehire_yes ~ 1)

wdata <- subset(sa_data, rehire == 0)
sa_surv_rehire_no <- Surv(wdata$tenure_cont, wdata$attrited)
sa_rehire_no_km_fit <- survfit(sa_surv_rehire_no ~ 1)


plot(sa_rehire_yes_km_fit$surv,
     main = 'Sales Associate Survival Estimate by Rehire Status', 
     ylab = "Survival Rate", xlab = "Days",
     lty = 2, pch = 20, col = "violetred")
points(sa_rehire_no_km_fit$surv, col = "turquoise4",lty = 2, pch = 20)
legend(200, 0.99, legend = c('Rehire - Yes','Rehire - No'), 
       col = c("violetred","turquoise4"),
       lwd = c(3,3), pch = 20)
```

Highest education:

```{r}
wdata <- subset(sa_data, high_school == 1)
sa_surv_hs <- Surv(wdata$tenure_cont, wdata$attrited)
sa_hs_km_fit <- survfit(sa_surv_hs ~ 1)

wdata <- subset(sa_data, associates == 1)
sa_surv_associates <- Surv(wdata$tenure_cont, wdata$attrited)
sa_associates_km_fit <- survfit(sa_surv_associates ~ 1)

wdata <- subset(sa_data, bachelor == 1)
sa_surv_bachelor <- Surv(wdata$tenure_cont, wdata$attrited)
sa_bachelor_km_fit <- survfit(sa_surv_bachelor ~ 1)

wdata <- subset(sa_data, masters_plus == 1)
sa_surv_masters_plus <- Surv(wdata$tenure_cont, wdata$attrited)
sa_masters_plus_km_fit <- survfit(sa_surv_masters_plus ~ 1)

wdata <- subset(sa_data, college == 1)
sa_surv_college_yes <- Surv(wdata$tenure_cont, wdata$attrited)
sa_college_yes_km_fit <- survfit(sa_surv_college_yes ~ 1)

wdata <- subset(sa_data, college == 0)
sa_surv_college_no <- Surv(wdata$tenure_cont, wdata$attrited)
sa_college_no_km_fit <- survfit(sa_surv_college_no ~ 1)


plot(sa_hs_km_fit$surv,
     main = 'Education Level', 
     ylab = "%", xlab = "Days",
     lty = 2, pch = 20, col = "slateblue2", xlim = c(0,730))
points(sa_associates_km_fit$surv, col = "darkolivegreen",lty = 2, pch = 20)
points(sa_bachelor_km_fit$surv, col = "violetred4",lty = 2, pch = 20)
points(sa_masters_plus_km_fit$surv, col = "yellow4",lty = 2, pch = 20)
legend(200, 0.99, legend = c('High School','Associates','Bachelors','Masters+'), 
       col = c("slateblue2","darkolivegreen",'violetred4','yellow4'),
       lwd = c(3,3), pch = 20)

```

Survival by Hours Available:

```{r}
wdata <- subset(sa_data, weekends_only == 1)
sa_surv_weekend <- Surv(wdata$tenure_cont, wdata$attrited)
sa_weekend_km_fit <- survfit(sa_surv_weekend ~ 1)


wdata <- subset(sa_data, weekdays_only == 1)
sa_surv_weekday <- Surv(wdata$tenure_cont, wdata$attrited)
sa_weekday_km_fit <- survfit(sa_surv_weekday ~ 1)


wdata <- subset(sa_data, both == 1)
sa_surv_both <- Surv(wdata$tenure_cont, wdata$attrited)
sa_both_km_fit <- survfit(sa_surv_both ~ 1)



plot(sa_both_km_fit$surv,
     main = 'Time Available',
     ylab = "Survival Rate", xlab = "Days",
     lty = 2, pch = 20)
points(sa_weekend_km_fit$surv, col = "turquoise",lty = 2, pch = 20)
points(sa_weekday_km_fit$surv, col = "salmon",lty = 2, pch = 20)

legend(400, 0.99, legend = c('Both','Weekend Only','Weekday Only'),
       col = c("black","turquoise","salmon"),
       lwd = c(3,3,3), pch = 20)

```

Sales Associate Survival Estimate by Assessment Score Group:

```{r}
#check for NA values
#sum(is.na(sa_data_sa_app$overall_percentile))

#Create catagorical assessment score feature
sa_data_sa_app$assessment_group <- ifelse(sa_data_sa_app$overall_percentile < 30, "Zero to <30",
                                          ifelse(sa_data_sa_app$overall_percentile < 70, "30 to <70","70 to 100"))

#Subsetting applicants with assessment_group == "Zero to <30"
wdata <- subset(sa_data_sa_app, assessment_group == "Zero to <30")
sa_surv_0_30 <- Surv(wdata$tenure_cont, wdata$attrited)
sa_0_30_km_fit <- survfit(sa_surv_0_30 ~ 1)

#Subsetting applicants with assessment_group == "30 to 70"
wdata <- subset(sa_data_sa_app, assessment_group == "30 to <70")
sa_surv_30_70 <- Surv(wdata$tenure_cont, wdata$attrited)
sa_30_70_km_fit <- survfit(sa_surv_30_70 ~ 1)

#Subsetting applicants with assessment_group == "70 to 100"
wdata <- subset(sa_data_sa_app, assessment_group == "70 to 100")
sa_surv_70_100 <- Surv(wdata$tenure_cont, wdata$attrited)
sa_70_100_km_fit <- survfit(sa_surv_70_100 ~ 1)


plot(sa_0_30_km_fit$surv,
     main = 'Assessment Score Group', 
     ylab = "Survival Rate", xlab = "Days",
     lty = 2, pch = 20)
points(sa_30_70_km_fit$surv, col = "blue", lty = 2, pch = 20)
points(sa_70_100_km_fit$surv, col = "orange", lty = 2, pch = 20)

legend(400, 0.99, legend = c('0 to 30','30 to 70','70 to 100'), 
       col = c("black","blue","orange"),
       lwd = c(3,3,3),pch = 20)
```

Store or Regional Information:Create Regional Identifiers

```{r, include=FALSE}
sa_data$canada_flag <- NA
canada_provinces <- "(Northwest Territories|Nunavut|Ontario|Prince Edward Island|Yukon|New Brunswick|Manitoba|Newfoundland|Nova Scotia|Saskatchewan|Quebec|British Columbia|Alberta)"
sa_data$canada_flag[grepl(canada_provinces, sa_data$home_state_province) == TRUE] <- 1
sa_data$canada_flag[grepl(canada_provinces, sa_data$home_state_province) == FALSE] <- 0

#Regional compariosn
wdata <- subset(sa_data, canada_flag == 1)
sa_surv_canada <- Surv(wdata$tenure_cont, wdata$attrited)
sa_canada_km_fit <- survfit(sa_surv_canada ~ 1)

wdata <- subset(sa_data, canada_flag == 0)
sa_surv_us <- Surv(wdata$tenure_cont, wdata$attrited)
sa_us_km_fit <- survfit(sa_surv_us ~ 1)
```

```{r}
plot(sa_us_km_fit$surv,
     main = 'By Region', 
     ylab = "Survival Rate", xlab = "Days",
     lty = 2, pch = 20, col = "blue")
points(sa_canada_km_fit$surv, col = "red",lty = 2, pch = 20)
legend(400, 0.99, legend = c('Region A','Region B'), 
       col = c("blue","red"),
       lwd = c(3,3), pch = 20)
```



## Part 5: Cox Proportional Hazard Regression Model

```{r}

#Create binary assessment score features
sa_data$assess_0_30 <- ifelse(sa_data$overall_percentile < 30, 1,0)
sa_data$assess_30_70 <- ifelse(sa_data$overall_percentile >= 30 & sa_data$overall_percentile < 70, 1,0)
sa_data$assess_70_100 <- ifelse(sa_data$overall_percentile >= 70, 1,0)
```
```{r}
#Binarize 'job applied for' feature
sa_data$sa_app <- ifelse(sa_data$new_job_group_application == "Retail Salespersons", 1,0)
sa_data$sl_app <- ifelse(sa_data$new_job_group_application == "First-Line Supervisors of Retail Sales Workers", 1,0)
```
```{r}
#Initializing time availability binary features
sa_data$time_slots_available <- 0
sa_data$weekends_only <- NA
sa_data$weekdays_only <- NA
sa_data$both <- NA

#Populating the features with values
for (i in 1:nrow(sa_data)){
  temp <- as.character(sa_data$times_available[i])
  sa_data$time_slots_available[i] <- length(unlist(strsplit(temp,";")))
  sa_data$weekends_only[i] <- ifelse(grepl("(Sat|Sun)",temp) == TRUE & grepl("(Mon|Tues|Wed|Thurs|Fri)",temp) == FALSE,1,0)
  sa_data$weekdays_only[i] <- ifelse(grepl("(Sat|Sun)",temp) == FALSE & grepl("(Mon|Tues|Wed|Thurs|Fri)",temp) == TRUE,1,0)
  sa_data$both[i] <- ifelse(grepl("(Sat|Sun)",temp) == TRUE & grepl("(Mon|Tues|Wed|Thurs|Fri)",temp) == TRUE,1,0)
}

#Fill in NA if no time slots selected
sa_data$weekends_only[sa_data$time_slots_available == 0] <- NA
sa_data$weekdays_only[sa_data$time_slots_available == 0] <- NA
sa_data$both[sa_data$time_slots_available == 0] <- NA
```

```{r include=FALSE}
##########################################
#KAPLAN MEIER ESTIMATES #################
#Build the standard survival object.
#tenure_cont records survival time (time to event)
#attrited indicates whether the applicant's departure was observed (status = 1) or that survival time was censored (status = 0). 
#a "+" after the time in the print out of KM indicates censoring.
#sa_surv <- Surv(sa_data$tenure_cont, sa_data$attrited)
#sa_surv
#Kaplan-Meier estimates of the rate of survival over time.
#sa_km_fit <- survfit(sa_surv ~ 1)
```


# Computing Cox PH Model

Round 1: Fitting the COX PH model to associate survival hazard with potential predictors

```{r}
set.seed(22)

#Create the formula
form <- formula(sa_surv ~ assess_0_30 + assess_30_70 + assess_70_100 + 
                  weekends_only + weekdays_only + both + 
                  sa_app + sl_app + rehire + 
                high_school + associates + bachelor + masters_plus)
```
```{r}
#Fit Cox Model
sa_cox <- coxph(form, data = sa_data,x = TRUE)
summary(sa_cox)
```

Plot survival estimates based on the model:
```{r}
sa_cox_fit <- survfit(sa_cox)
plot(sa_km_fit$surv,
     main = 'Sales Associate Survival Estimate', ylab = "Survival Rate", xlab = "Days",
     xlim = c(0,365), lty = 2, pch = 20)
points(sa_cox_fit$surv, col = "lightblue",lty = 2, pch = 20)
legend(200, 0.99, legend = c('All Hires','Cox PH'), 
       col = c("black","lightblue"), pch = c(20,20),
       lwd = c(2,2))
```

Round 2: Kept only  statistically significant features at 5% significance value (dropped 'weekdays_only', 
'both', 'assess_70_100','sl_app' and 'masters_plus')
```{r}
form <- formula(sa_surv ~ assess_0_30 + assess_30_70 + 
                  weekends_only +
                  sa_app + rehire + 
                  high_school + associates + bachelor)

#Fit Cox Model
sa_cox <- coxph(form, data = sa_data,x = TRUE)
summary(sa_cox)

#Plot survival estimates
sa_cox_fit <- survfit(sa_cox)
plot(sa_km_fit$surv,
     main = 'Sales Associate Survival Estimate', ylab = "Survival Rate", xlab = "Days",
     xlim = c(0,365), lty = 2, pch = 20)
points(sa_cox_fit$surv, col = "lightblue",lty = 2, pch = 20)
legend(200, 0.99, legend = c('All Hires','Cox PH'), 
       col = c("black","lightblue"), pch = c(20,20),
       lwd = c(2,2))
```


Extracting predicted survival rates

```{r}
newdata <- subset(sa_data, select = c(applicant_id, assess_0_30,assess_30_70,assess_70_100,
                                      weekends_only,weekdays_only,both,
                                      sa_app,sl_app, rehire, college,
                                    high_school, associates,bachelor, masters_plus))
#Base hazard at time t
bh=basehaz(sa_cox)

#Example specific points in time (days) 
times <- c(10,30,90,180,365,730)

#Sample applicant IDs
IDs<-c(14704065,14740260,15014188,15020641,15091641)

#Subset data only for these IDs
Applicants_select= subset(sa_data, sa_data$applicant_id %in% IDs)

#Create empty columsn for the specified points in time
Applicants_select$T_10<-NA
Applicants_select$T_30<-NA
Applicants_select$T_90<-NA
Applicants_select$T_180<-NA
Applicants_select$T_365<-NA
Applicants_select$T_730<-NA


#Formula to calculate Survival at a given point in time for a given person p:
#exp(-bh$hazard[times])^(exp(sum(sa_cox$coefficients*newdata[person,], na.rm = TRUE)))
for (i in 1: length(IDs)){
  
    Applicants_select$T_10[i]=exp(-bh$hazard[10])^(exp(sum(sa_cox$coefficients*newdata[newdata$applicant_id==IDs[i],][, -(which(names(newdata) %in% "applicant_id"))], na.rm = TRUE)))
    
    Applicants_select$T_30[i]=exp(-bh$hazard[30])^(exp(sum(sa_cox$coefficients*newdata[newdata$applicant_id==IDs[i],][, -(which(names(newdata) %in% "applicant_id"))], na.rm = TRUE)))
    
    Applicants_select$T_90[i]=exp(-bh$hazard[90])^(exp(sum(sa_cox$coefficients*newdata[newdata$applicant_id==IDs[i],][, -(which(names(newdata) %in% "applicant_id"))], na.rm = TRUE)))
    
    Applicants_select$T_180[i]=exp(-bh$hazard[180])^(exp(sum(sa_cox$coefficients*newdata[newdata$applicant_id==IDs[i],][, -(which(names(newdata) %in% "applicant_id"))], na.rm = TRUE)))
    
    Applicants_select$T_365[i]=exp(-bh$hazard[365])^(exp(sum(sa_cox$coefficients*newdata[newdata$applicant_id==IDs[i],][, -(which(names(newdata) %in% "applicant_id"))], na.rm = TRUE)))
    
    Applicants_select$T_730[i]=exp(-bh$hazard[730])^(exp(sum(sa_cox$coefficients*newdata[newdata$applicant_id==IDs[i],][, -(which(names(newdata) %in% "applicant_id"))], na.rm = TRUE)))
  
}

#Saving survival rates at the specified points in time for the randomly selected applicants
#Applicants_select=subset(Applicants_select,select = c("T_10","T_30","T_90","T_180","T_365","T_730"))
#write.csv(Applicants_select, file = "Applicants_select.csv")
```


# Model Diagnostics

```{r}
#Testing  for the proportional-hazards (PH) assumption
#A p-value less than 0.05 indicates a violation of the proportionality assumption.
test.ph <- cox.zph(sa_cox)
test.ph


#Identfying outliers
dresids <- residuals( sa_cox, type="deviance" )
lp <- predict(sa_cox, type="lp" )
plot(lp, dresids, xlab="Linear Predictor", ylab="Deviance Residual")
########################

#Testing for influential observations
#ggcoxdiagnostics(sa_cox, type = , linear.predictions = TRUE)
#dfbetas <- residuals(sa_cox, type="dfbetas")
